---
title: "data processing - neighborhood social ties"
output: html_document
date: "2024-01-17"
---

# Neighborhood social ties scale
# From Kuo et al. 1998
https://link.springer.com/article/10.1023/A:1022294028903

The central dependent variable in this study was Neighborhood Social Ties (NSTs). NSTs were assessed through 8 items with responses on a 5-point Likert scale from 0 (not at all) to 4 (very much). 
Principal components
analysis with varimax rotation yielded three factors; items for the factors were summed to create three subscales, each of which demonstrated acceptable reliability
 1. Socializing includes two items 
      ("Do you have many visitors every day?" and "Do you socialize a lot within the building?")
 2. Nearby Neighbors includes two items 
      ("How well do you know the people next door?" and "How well do you know the people on your floor?") 
 3. Local Sense of Community includes three items 
      ("Are people here concerned with helping and supporting one another?", "Do people here acknowledge one another when passing in the hallway?", and "Is there a strong feeling of belonging here?")
      
Notes: We seem to be missing one item from the scale since we only asked 7 questions (the question "Do neighbors help each other out by lending things like tools, giving someone a ride, or watching each other's houses when they're away?" from the original survey is missing)

In the original paper the overall scales and subscales seem to be averaged


## Examine raw inputs (check for weird and missing things)
```{r NST-examine-raw-inputs}
# Baseline
tbl_summary(
  baseline, 
  include = c(448:454),
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```


```{r NST-examine-raw-inputs}
# w2
tbl_summary(
  w2, 
  include = c(315:321),
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```

```{r NST-examine-raw-inputs}
# w3
tbl_summary(
  w3, 
  include = c(295:301),
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```

```{r construct-NST-scale}
#Baseline
baseline_scales <- baseline %>% 
  mutate(NST_item1 = `@74673`,
         NST_item2 = `@35201`,
         NST_item3 = `@53819`,
         NST_item4 = `@51902`,
         NST_item5 = `@49388`,
         NST_item6 = `@11637`,
         NST_item7 = `@78709`) 

# Construct NST score as arithmetic mean of individual items 
# If three items or fewer are missing, then still construct the scale
baseline_scales <- baseline_scales %>% 
     rowwise() %>% 
     mutate(count_NST_nas = sum(is.na(c_across(starts_with("NST_")))),
       NST_score_baseline = case_when(
         count_NST_nas <=3 ~ mean(c_across(starts_with("NST_")), na.rm = TRUE),
         count_NST_nas >3 ~ NA)) %>% 
     ungroup() 
```

```{r construct-NST-scale}
#w2
w2_scales <- w2 %>% 
  mutate(NST_item1 = `@74673`,
         NST_item2 = `@35201`,
         NST_item3 = `@53819`,
         NST_item4 = `@51902`,
         NST_item5 = `@49388`,
         NST_item6 = `@11637`,
         NST_item7 = `@78709`) 


# Construct NST score as arithmetic mean of individual items 
# If three items or fewer are missing, then still construct the scale
w2_scales <- w2_scales %>% 
     rowwise() %>% 
     mutate(count_NST_nas = sum(is.na(c_across(starts_with("NST_")))),
       NST_score_w2 = case_when(
         count_NST_nas <=3 ~ mean(c_across(starts_with("NST_")), na.rm = TRUE),
         count_NST_nas >3 ~ NA)) %>% 
     ungroup() 
```

```{r construct-NST-scale}
#w3
w3_scales <- w3 %>% 
  mutate(NST_item1 = `@74673`,
         NST_item2 = `@35201`,
         NST_item3 = `@53819`,
         NST_item4 = `@51902`,
         NST_item5 = `@49388`,
         NST_item6 = `@11637`,
         NST_item7 = `@78709`) 

# Construct NST score as arithmetic mean of individual items 
# If three items or fewer are missing, then still construct the scale
w3_scales <- w3_scales %>% 
     rowwise() %>% 
     mutate(count_NST_nas = sum(is.na(c_across(starts_with("NST_")))),
       NST_score_w3 = case_when(
         count_NST_nas <=3 ~ mean(c_across(starts_with("NST_")), na.rm = TRUE),
         count_NST_nas >3 ~ NA)) %>% 
     ungroup() 
```

```{r}
# Check
baseline_scales %>% 
  select(448:454, NST_score_baseline, NST.Total) 
```
```{r}
# Check
w2_scales %>% 
  select(315:321, NST_score_w2, NST.Total) 
```


# Simplify datasets to needed variables
```{r}
baseline_subset <- baseline_scales %>% 
  select(ID, Cohort, NST_score_baseline, Group, SITE.NUMBER)

w2_subset <- w2_scales %>% 
  select(ID, NST_score_w2)

w3_subset <- w3_scales %>% 
  select(ID, NST_score_w3)
```


# Combine w2 and w3
```{r}
w2w3_combined <- w2_subset %>% 
  full_join(w3_subset, by = "ID") %>% 
  mutate(NST_score_w2w3 = case_when(
    !is.na(NST_score_w2) ~ NST_score_w2,
    !is.na(NST_score_w3) ~ NST_score_w3,
    TRUE ~ NA
  )) %>% 
  select(ID, NST_score_w2w3)
```

# Join w2w3 to baseline
# Left join so that supplemental people without baseline are excluded
```{r}
long_data <- baseline_subset %>% 
  left_join(w2w3_combined, by = "ID")
```


# Subset to longitudinal participants with baseline and follow-up for SCT
```{r}
NST_complete <- long_data %>% 
  mutate(missing_baseline = is.na(NST_score_baseline),
         missing_followup = is.na(NST_score_w2w3)) %>% 
  filter(missing_baseline == FALSE & missing_followup == FALSE) %>% 
  select(-missing_baseline, -missing_followup)
```

# Pivot data so there is only one y column
```{r}
NST_complete_pivot <- NST_complete %>% 
  pivot_longer(cols = c(NST_score_baseline, NST_score_w2w3), names_to = "time", 
               names_prefix = "NST_score_", values_to = "NST_score")
  
```



# View distribution of outcome by group and wave
```{r}
NST_complete_pivot %>% 
  ggplot(mapping = aes(x = NST_score)) +
  geom_histogram(bins = 15) +
  facet_grid(vars(Group), vars(time))
```

# Manual diff-in-diff to plot
```{r, fig.width = 8, fig.height = 4}
NST_plot_vals <- NST_complete_pivot %>% 
  mutate(group = case_when(
    Group == 1 ~ "Intervention",
    Group == 0 ~ "Control"
  )) %>% 
  group_by(group, time) %>% 
  summarise(mean_NST = mean(NST_score),
            se_NST = sd(NST_score) / sqrt(n()),
            upper = mean_NST +(-1.96 * se_NST),
            lower = mean_NST + (1.96 * se_NST))

NST_plot_vals %>% 
  ggplot(mapping = aes(x = time, y = mean_NST, color = group)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), fatten = 8, linewidth = 1) +
  geom_line(aes(group = group), size = 1) 
```

