---
title: "Data Processing - SCI 2"
author: "Rachel L. Thompson"
date: "2023-12-01"
output: html_document
---

```{r load-packages}
library(tidyverse)
library(haven)
library(gtsummary)
library(lme4)
library(PerformanceAnalytics)
```


# Read in SPSS data
```{r read-data, message=FALSE, include=FALSE}
baseline <- read_sav("Data/MASTER baseline _Adult and Youth merged FINAL 7.5.20.sav")

w2 <- read_sav("Data/MASTER w2 adult and youth merged FIXED.sav")

w3 <- read_sav("Data/MASTER W3 adult and youth merged FINAL.sav")

```


## Manually Recode @87354 for W2 Supplements with 0s (Should be 5)
```{r}
# w2_fixed <- w2 %>%
#   mutate(fixed_87354 = case_when(
#     ID == 112064 ~ 5,
#     ID == 116064 ~ 5,
#     ID == 111061 ~ 5,
#     ID == 105062 ~ 5,
#     ID == 118064 ~ 5,
#     ID == 117073 ~ 5,
#     ID == 207065 ~ 5,
#     ID == 112073 ~ 5,
#     ID == 108023 ~ 5,
#     ID == 108070 ~ 5,
#     ID == 111067 ~ 5,
#     TRUE ~ `@87354`
#   ))
# 
# write_sav(w2_fixed, path = "Data/MASTER w2 adult and youth merged FIXED.sav")
```




```{r}
baseline_doubles <- baseline %>% 
  group_by(ID) %>% 
  summarise(n = n()) 

baseline_doubles %>% 
  filter(n > 1) 
```



```{r}
w2_doubles <- w2 %>% 
  mutate(ID_num = as.numeric(ID)) %>% 
  group_by(ID_num) %>% 
  summarise(n = n()) 

w2_dup_IDs <- w2_doubles %>% 
  filter(n > 1) %>% 
  select(ID_num) #%>% 
  # left_join(w2, by = "ID")

w2_dup_IDs <- as.list(w2_dup_IDs$ID_num)

w2 <- w2 %>% 
  filter(!(as.numeric(ID) %in% w2_dup_IDs)) 
```



```{r}
w3_doubles <- w3 %>% 
  group_by(ID) %>% 
  summarise(n = n()) 

w3_doubles %>% 
  filter(n > 1)
```




# Overall sample size check
```{r}
baseline %>% 
  group_by(Cohort) %>% 
  summarise(n = n())
```


```{r}
w2 %>% 
  group_by(COHORT) %>% 
  summarise(n = n())
```


```{r}
w3 %>% 
  group_by(COHORT) %>% 
  summarise(n = n())
```

# Construct Scales at Each Time Point

## Sense of Community Index II

https://dl.icdst.org/pdfs/files/f458f0f15016819295377e5a979b1893.pdf 

Chavis, D.M., Lee, K.S., & Acosta J.D. (2008). The Sense of Community (SCI) Revised: The Reliability and
Validity of the SCI-2. Paper presented at the 2nd International Community Psychology Conference,
Lisboa, Portugal.

###Instructions for Scoring the Revised Sense of Community Index
#### Identifying the Community Referent  
The attached scale was developed to be used in many different types of communities. Be sure to specify
the type of community the scale is referring to before administering the scale. Do not use “your
community” as the referent.  

#### Interpreting the Initial Question  
The initial question “How important is it to you to feel a sense of community with other community
members?” is a validating question that can be used to help you interpret the results. We have found
that total sense of community is correlated with this question – but keep in mind this may not be true in
every community.

#### Scoring the Scale
For the 24 questions that comprise the revised Sense of Community Index participants:

- Not at All = 0, Somewhat = 1, Mostly = 2, Completely = 3
- Total Sense of Community Index = Sum of Q1 to Q24

Subscales 
- Reinforcement of Needs = Q1 + Q2 + Q3 + Q4 + Q5 + Q6
- Membership = Q7 + Q8 + Q9 + Q10 + Q11 + Q12
- Influence = Q13 + Q14 + Q15 + Q16 + Q17 + Q18
- Shared Emotional Connection = Q19 + Q20 + Q21 + Q22 + Q23 + Q24

## Examine Raw Inputs (check for weird and missing things)
```{r}
# Baseline
tbl_summary(
  baseline,
  include = c(271:295), 
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```

```{r}
# W2
tbl_summary(
  w2,
  include = c(149:173), 
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```

```{r}
# W3
tbl_summary(
  w3,
  include = c(187:211), 
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```


# Construct SCI II and Subscales
```{r}
# Construct SCI II Score from questions - baseline
baseline_scales <- baseline %>% 
    rowwise() %>% 
    mutate(SCI2_score_baseline = sum(c_across(272:295)),
           count_nas = sum(is.na(c_across(272:295))),
           AVG_SCI2_score_baseline = case_when(
             count_nas <= 5 ~ mean(c_across(272:295), na.rm = TRUE),
             count_nas > 5 ~ NA
           )) %>% 
    ungroup()
```


```{r}
# Construct SCI II Score from questions - w2
w2_scales <- w2 %>% 
    rowwise() %>% 
    mutate(SCI2_score_w2 = sum(c_across(150:173)),
           count_nas = sum(is.na(c_across(150:173))),
           AVG_SCI2_score_w2 = case_when(
             count_nas <= 5 ~ mean(c_across(150:173), na.rm = TRUE),
             count_nas > 5 ~ NA
           )) %>% 
    ungroup()
```

```{r}
# Construct SCI II Score from questions - w3
w3_scales <- w3 %>% 
    rowwise() %>% 
    mutate(SCI2_score_w3 = sum(c_across(188:211)),
           count_nas = sum(is.na(c_across(188:211))),
           AVG_SCI2_score_w3 = case_when(
             count_nas <= 5 ~ mean(c_across(188:211), na.rm = TRUE),
             count_nas > 5 ~ NA
           )) %>% 
    ungroup()
```

```{r}
# Check
baseline_scales %>% 
  select(272:295, SCI2_score_baseline, SCI.II.Total, AVG_SCI2_score_baseline) 
```

```{r}
# Check
w2_scales %>% 
  select(150:173, SCI2_score_w2, SCI.II.Total, AVG_SCI2_score_w2) 
```

```{r}
# Check
w3_scales %>% 
  select(188:211, SCI2_score_w3, SCI.II.Total, AVG_SCI2_score_w3) 
```


```{r}
# Construct SCI 2 subscales - baseline
baseline_scales <- baseline_scales %>% 
  mutate(SCI2_reinforcement_of_needs_baseline = `@46242` + `@42897` + `@65379` + `@71718` + 
           `@43987` + `@38496`,
         SCI2_membership_baseline = `@24315` + `@80714` + `@24359` + `@63377` + `@89515` + `@39920`,
         SCI2_influence_baseline = `@15989` + `@12255` + `@87666` + `@31253` + `@92328` + `@20836`, 
         SCI2_shared_emotional_connection_baseline = `@77731`+ `@25861` + `@60118`+ `@16393` + `@87417` +
           `@18861`) 
```


```{r}
# Construct SCI 2 subscales - w2
w2_scales <- w2_scales %>% 
  mutate(SCI2_reinforcement_of_needs_w2 = `@46242` + `@42897` + `@65379` + `@71718` + 
           `@43987` + `@38496`,
         SCI2_membership_w2 = `@24315` + `@80714` + `@24359` + `@63377` + `@89515` + `@39920`,
         SCI2_influence_w2 = `@15989` + `@12255` + `@87666` + `@31253` + `@92328` + `@20836`, 
         SCI2_shared_emotional_connection_w2 = `@77731`+ `@25861` + `@60118`+ `@16393` + `@87417` +
           `@18861`) 
```

```{r}
# Construct SCI 2 subscales - w3
w3_scales <- w3_scales %>% 
  mutate(SCI2_reinforcement_of_needs_w3 = `@46242` + `@42897` + `@65379` + `@71718` + 
           `@43987` + `@38496`,
         SCI2_membership_w3 = `@24315` + `@80714` + `@24359` + `@63377` + `@89515` + `@39920`,
         SCI2_influence_w3 = `@15989` + `@12255` + `@87666` + `@31253` + `@92328` + `@20836`, 
         SCI2_shared_emotional_connection_w3 = `@77731`+ `@25861` + `@60118`+ `@16393` + `@87417` +
           `@18861`) 
```


```{r}
# Check
baseline_scales %>% 
  select(starts_with("SCI2"), starts_with("SCI.II"))
```

```{r}
# Check
w2_scales %>% 
  select(starts_with("SCI2"), starts_with("SCI.II"))
```

```{r}
# Check
w3_scales %>% 
  select(starts_with("SCI2"), starts_with("SCI.II"))
```

# SCI 2 Score vs. Validating Question
```{r}
# Baseline
tbl_summary(
  baseline_scales, 
  include = c(SCI2_score_baseline, AVG_SCI2_score_baseline),
  by = `@87354`
)
```

```{r}
# w2
tbl_summary(
  w2_scales, 
  include = c(SCI2_score_w2, AVG_SCI2_score_w2),
  by = fixed_87354
)
```

```{r}
# w3
tbl_summary(
  w3_scales, 
  include = c(SCI2_score_w3, AVG_SCI2_score_w3),
  by = `@87354`
)
```

## ANALYSIS 1: SCI2 Score as SUM
# Simplify datasets to needed variables
```{r}
baseline_subset <- baseline_scales %>% 
  select(ID, Cohort, SCI2_score_baseline, Group, SITE.NUMBER)

w2_subset <- w2_scales %>% 
  select(ID, SCI2_score_w2)

w3_subset <- w3_scales %>% 
  select(ID, SCI2_score_w3)
```

# Combine w2 and w3
```{r}
w2w3_combined <- w2_subset %>% 
  full_join(w3_subset, by = "ID") %>% 
  mutate(SCI2_score_w2w3 = case_when(
    !is.na(SCI2_score_w2) ~ SCI2_score_w2,
    !is.na(SCI2_score_w3) ~ SCI2_score_w3,
    TRUE ~ NA
  )) %>% 
  select(ID, SCI2_score_w2w3)
```

# Join w2w3 to baseline
# Left join so that supplemental people without baseline are excluded
```{r}
long_data <- baseline_subset %>% 
  left_join(w2w3_combined, by = "ID")
```



# Subset to longitudinal participants with baseline and follow-up for SCI-2
```{r}
SCI2_complete <- long_data %>% 
  mutate(missing_baseline = is.na(SCI2_score_baseline),
         missing_followup = is.na(SCI2_score_w2w3)) %>% 
  filter(missing_baseline == FALSE & missing_followup == FALSE) %>% 
  select(-missing_baseline, -missing_followup)
```

# Pivot data so there is only one y column
```{r}
SCI2_complete_pivot <- SCI2_complete %>% 
  pivot_longer(cols = c(SCI2_score_baseline, SCI2_score_w2w3), names_to = "time", 
               names_prefix = "SCI2_score_", values_to = "SCI2_score")
  
```
 
 
# View distribution of outcome by group and wave
```{r}
SCI2_complete_pivot %>% 
  ggplot(mapping = aes(x = SCI2_score)) +
  geom_histogram(bins = 15) +
  facet_grid(vars(Group), vars(time))
```


# Manual diff-in-diff to plot
```{r, fig.width = 8, fig.height = 4}
SCI2_plot_vals <- SCI2_complete_pivot %>% 
  mutate(group = case_when(
    Group == 1 ~ "Intervention",
    Group == 0 ~ "Control"
  )) %>% 
  group_by(group, time) %>% 
  summarise(mean_SCI2 = mean(SCI2_score),
            se_SCI2 = sd(SCI2_score) / sqrt(n()),
            upper = mean_SCI2 +(-1.96 * se_SCI2),
            lower = mean_SCI2 + (1.96 * se_SCI2))

SCI2_plot_vals %>% 
  ggplot(mapping = aes(x = time, y = mean_SCI2, color = group)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), fatten = 8, linewidth = 1) +
  geom_line(aes(group = group), size = 1) 
```





## ANALYSIS 2: SCI2 Score as AVG
# Simplify datasets to needed variables
```{r}
baseline_subset <- baseline_scales %>% 
  select(ID, Cohort, AVG_SCI2_score_baseline, Group, SITE.NUMBER)

w2_subset <- w2_scales %>% 
  select(ID, AVG_SCI2_score_w2)

w3_subset <- w3_scales %>% 
  select(ID, AVG_SCI2_score_w3)
```

# Combine w2 and w3
```{r}
w2w3_combined <- w2_subset %>% 
  full_join(w3_subset, by = "ID") %>% 
  mutate(AVG_SCI2_score_w2w3 = case_when(
    !is.na(AVG_SCI2_score_w2) ~ AVG_SCI2_score_w2,
    !is.na(AVG_SCI2_score_w3) ~ AVG_SCI2_score_w3,
    TRUE ~ NA
  )) %>% 
  select(ID, AVG_SCI2_score_w2w3)
```

# Join w2w3 to baseline
# Left join so that supplemental people without baseline are excluded
```{r}
long_data <- baseline_subset %>% 
  left_join(w2w3_combined, by = "ID")
```



# Subset to longitudinal participants with baseline and follow-up for SCI-2
```{r}
SCI2_complete <- long_data %>% 
  mutate(missing_baseline = is.na(AVG_SCI2_score_baseline),
         missing_followup = is.na(AVG_SCI2_score_w2w3)) %>% 
  filter(missing_baseline == FALSE & missing_followup == FALSE) %>% 
  select(-missing_baseline, -missing_followup)
```

# Pivot data so there is only one y column
```{r}
SCI2_complete_pivot <- SCI2_complete %>% 
  pivot_longer(cols = c(AVG_SCI2_score_baseline, AVG_SCI2_score_w2w3), names_to = "time", 
               names_prefix = "AVG_SCI2_score_", values_to = "AVG_SCI2_score")
  
```
 
 
# View distribution of outcome by group and wave
```{r}
SCI2_complete_pivot %>% 
  ggplot(mapping = aes(x = AVG_SCI2_score)) +
  geom_histogram(bins = 15) +
  facet_grid(vars(Group), vars(time))
```

# Manual diff-in-diff to plot
```{r, fig.width = 8, fig.height = 4}
SCI2_plot_vals <- SCI2_complete_pivot %>% 
  mutate(group = case_when(
    Group == 1 ~ "Intervention",
    Group == 0 ~ "Control"
  )) %>% 
  group_by(group, time) %>% 
  summarise(mean_SCI2 = mean(AVG_SCI2_score),
            se_SCI2 = sd(AVG_SCI2_score) / sqrt(n()),
            upper = mean_SCI2 +(-1.96 * se_SCI2),
            lower = mean_SCI2 + (1.96 * se_SCI2))

SCI2_plot_vals %>% 
  ggplot(mapping = aes(x = time, y = mean_SCI2, color = group)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), fatten = 8, linewidth = 1) +
  geom_line(aes(group = group), size = 1) 
```





