---
title: "data processing - social cohesion and trust"
output: html_document
date: "2024-01-17"
---

# Community Cohesion Scale- Social Cohesion and Trust Subscale
https://sparqtools.org/mobility-measure/collective-efficacy-scale/#all-survey-questions
https://docs.google.com/document/d/1UFKLMUbs-4VJ1-A_qmP__2OLyR6DFEW6JLkjy-RRm4g/edit

To complete the 5-item social cohesion and trust section, respondents use a different 5-point (1 = strongly agree; 5 = strongly disagree) to rate statements such as, “People around here are willing to help their neighbors.” Researchers later reverse-code the two negatively worded items so that high values mean more collective efficacy, and then calculate a score for each respondent by averaging their ratings.


## Examine raw inputs (check for weird and missing things)
```{r SCT-examine-raw-inputs}
# Baseline
tbl_summary(
  baseline, 
  include = c(386:390),
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```

```{r SCT-examine-raw-inputs}
# W2
tbl_summary(
  w2, 
  include = c(263:267),
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```

```{r SCT-examine-raw-inputs}
# W3
tbl_summary(
  w3, 
  include = c(316:320),
  missing = "always",
  type = all_continuous() ~ "categorical"
)
```



```{r construct-SCT-scale}
#Baseline
baseline_scales <- baseline %>% 
  # Recode variables to be on 1-5 point scale instead of 0-4
  # Reverse score items 4 and 5
  mutate(SCT_item1 = `@17958` + 1,
         SCT_item2 = `@23069` + 1,
         SCT_item3 = `@21767` + 1,
         SCT_item4 = (5 + 1) - (`@11280` + 1),
         SCT_item5 = (5 + 1) - (`@2409` + 1)) 

# Construct SCT score as arithmetic mean of individual items 
# If two items or fewer are missing, then still construct the scale
 baseline_scales <- baseline_scales %>% 
     rowwise() %>% 
     mutate(count_SCT_nas = sum(is.na(c_across(starts_with("SCT_")))),
       SCT_score_baseline = case_when(
         count_SCT_nas <=2 ~ mean(c_across(starts_with("SCT_")), na.rm = TRUE),
         count_SCT_nas >2 ~ NA)) %>% 
     ungroup() 
```

```{r construct-SCT-scale}
#w2
w2_scales <- w2 %>% 
  # Recode variables to be on 1-5 point scale instead of 0-4
  # Reverse score items 4 and 5
  mutate(SCT_item1 = `@17958` + 1,
         SCT_item2 = `@23069` + 1,
         SCT_item3 = `@21767` + 1,
         SCT_item4 = (5 + 1) - (`@11280` + 1),
         SCT_item5 = (5 + 1) - (`@2409` + 1)) 

# Construct SCT score as arithmetic mean of individual items 
# If two items or fewer are missing, then still construct the scale
 w2_scales <- w2_scales %>% 
     rowwise() %>% 
     mutate(count_SCT_nas = sum(is.na(c_across(starts_with("SCT_")))),
       SCT_score_w2 = case_when(
         count_SCT_nas <=2 ~ mean(c_across(starts_with("SCT_")), na.rm = TRUE),
         count_SCT_nas >2 ~ NA)) %>% 
     ungroup()
```

```{r construct-SCT-scale}
#w3
w3_scales <- w3 %>% 
  # Recode variables to be on 1-5 point scale instead of 0-4
  # Reverse score items 4 and 5
  mutate(SCT_item1 = `@17958` + 1,
         SCT_item2 = `@23069` + 1,
         SCT_item3 = `@21767` + 1,
         SCT_item4 = (5 + 1) - (`@11280` + 1),
         SCT_item5 = (5 + 1) - (`@2409` + 1)) 

# Construct SCT score as arithmetic mean of individual items 
# If two items or fewer are missing, then still construct the scale
 w3_scales <- w3_scales %>% 
     rowwise() %>% 
     mutate(count_SCT_nas = sum(is.na(c_across(starts_with("SCT_")))),
       SCT_score_w3 = case_when(
         count_SCT_nas <=2 ~ mean(c_across(starts_with("SCT_")), na.rm = TRUE),
         count_SCT_nas >2 ~ NA)) %>% 
     ungroup()
```

```{r}
# Check
baseline_scales %>% 
  select(386:390, SCT_score_baseline, SocialCohTrust.Total) 
```
```{r}
# Check
w2_scales %>% 
  select(263:267, SCT_score_w2, SocialCohTrust.Total) 
```

# Simplify datasets to needed variables
```{r}
baseline_subset <- baseline_scales %>% 
  select(ID, Cohort, SCT_score_baseline, Group, SITE.NUMBER)

w2_subset <- w2_scales %>% 
  select(ID, SCT_score_w2)

w3_subset <- w3_scales %>% 
  select(ID, SCT_score_w3)
```


# Combine w2 and w3
```{r}
w2w3_combined <- w2_subset %>% 
  full_join(w3_subset, by = "ID") %>% 
  mutate(SCT_score_w2w3 = case_when(
    !is.na(SCT_score_w2) ~ SCT_score_w2,
    !is.na(SCT_score_w3) ~ SCT_score_w3,
    TRUE ~ NA
  )) %>% 
  select(ID, SCT_score_w2w3)
```

# Join w2w3 to baseline
# Left join so that supplemental people without baseline are excluded
```{r}
long_data <- baseline_subset %>% 
  left_join(w2w3_combined, by = "ID")
```


# Subset to longitudinal participants with baseline and follow-up for SCT
```{r}
SCT_complete <- long_data %>% 
  mutate(missing_baseline = is.na(SCT_score_baseline),
         missing_followup = is.na(SCT_score_w2w3)) %>% 
  filter(missing_baseline == FALSE & missing_followup == FALSE) %>% 
  select(-missing_baseline, -missing_followup)
```

# Pivot data so there is only one y column
```{r}
SCT_complete_pivot <- SCT_complete %>% 
  pivot_longer(cols = c(SCT_score_baseline, SCT_score_w2w3), names_to = "time", 
               names_prefix = "SCT_score_", values_to = "SCT_score")
  
```



# View distribution of outcome by group and wave
```{r}
SCT_complete_pivot %>% 
  ggplot(mapping = aes(x = SCT_score)) +
  geom_histogram(bins = 15) +
  facet_grid(vars(Group), vars(time))
```

# Manual diff-in-diff to plot
```{r, fig.width = 8, fig.height = 4}
SCT_plot_vals <- SCT_complete_pivot %>% 
  mutate(group = case_when(
    Group == 1 ~ "Intervention",
    Group == 0 ~ "Control"
  )) %>% 
  group_by(group, time) %>% 
  summarise(mean_SCT = mean(SCT_score),
            se_SCT = sd(SCT_score) / sqrt(n()),
            upper = mean_SCT +(-1.96 * se_SCT),
            lower = mean_SCT + (1.96 * se_SCT))

SCT_plot_vals %>% 
  ggplot(mapping = aes(x = time, y = mean_SCT, color = group)) +
  geom_pointrange(aes(ymin = lower, ymax = upper), fatten = 8, linewidth = 1) +
  geom_line(aes(group = group), size = 1) 
```
